### 2021-11-26 main.tex
Update on Overleaf.

\documentclass{article}
\usepackage[utf8]{inputenc}
%\usepackage[margin=2cm]{geometry}

\title{Tradeoffs between sustainable development goals in systems of cities}

\author{J. Raimbault$^{1,\ast}$ and D. Pumain$^{2}$\medskip\\
$^{1}$ CASA, University College London\\
$^{2}$ G{\'e}ographie-cit{\'e}s, Universit{\'e} Paris 1\medskip\\
$\ast$ \texttt{juste.raimbault@polytechnnique.edu}
}
\date{}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

The concept of

\section{Urban system model}

\section{Results}

\section{Discussion}

\bibliographystyle{apalike}
\bibliography{biblio}

\end{document}